The Essence of Summarization\\ in Pushdown Flow Analyses\\ \textbf{Supplemental Material}

### 2013-12-13 content.tex
Got down to page limit. Cut the proof of non-termination in problematic semantics. FIAT SEMANTICS

\paragraph{Example:} if we use this semantics to run through the example of the previous section, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either, so if the operator weren't $\le$ but instead $<$ or $+$, we could still constant fold that away.
%%

%%
The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics.
%
We follow up with a similar proof to that of PDCFA, which is that the CES$\xi$K machine has a bisimulation with its tabular semantics, with comparable allocation strategies.
\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
  The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
\end{theorem}
This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
%
%That is, if we always allocated bindings in the heap.
%
%The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
%
%The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
%
%With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
%

@@

%% TODO: Cut here next.
There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive, since it captures the entire stack, leading to space leaks \citep{ianjohnson:kiselyov:against-callcc}.
%
Alternative control operators have been proposed that delimit how much of the stack to capture, such as $\%$ (read ``prompt'') and capture operator ${\mathcal F}$ (read ``control'') \citep{ianjohnson:felleisen:control:1988}, or \texttt{reset} (equivalent to $\%$) and \texttt{shift} \citep{ianjohnson:danvy:filinski:delim:1990}.
%
However, the stacks captured by these operators always extend the stack when invoked, rather than replace it like those captured with \rackett{call/cc}.
%
Continuations that have this extension behavior are called ``composable continuations.''
%
Stack replacement is easily modeled in a regular analysis using the AAM approach, and Vardoulakis and Shivers showed it can be done in a pushdown approach.
%
Stack extension, however, poses a new challenge for pushdown analysis, since one application of a composable continuation means pushing an unbounded amount of frames onto the stack.
%
Vardoulakis' and Shivers' approach does not immediately apply in this situation, since their technique drops all knowledge of the stack at a continuation's invocation site; extension, however, must preserve it.
%%

@@

Turning this into a table-based semantics involves making prompts a point of indirection, just like function calls.
%
Memoization also gets a new context to consider, calling a continuation, because composable continuations act like functions.
%
The new semantic spaces are then

@@

\autoref{fig:shift-reset-table0} has what one might naturally write using just this information.
%
Unfortunately, since $\rt$ continuations contain a store, and continuations can now appear in the store, this introduces an unbounded data structure that could cause the analysis to never terminate.
%
For example, a loop that continually captures continuations would introduce unboundedly many continuation values due to the ever-growing store.
%
This means the store is not finite height, and there may be no fixed point.

\paragraph{A fix to non-termination:} we allow for a tunable approximation by using $\alloc$.
%
Captured continuations' store component in the $\rt$ context gets replaced with an address of possible stores that it would approximate.

@@

\section{Conclusion and future work}

>>>>

\paragraph{Example:} revising the earlier example, we find that with the addition of stack frames, there is no confusion about \texttt{z}'s value either.
%
If the operator were instead $<$ or $+$, we could still constant fold that away.
%%

%%
The stack frame introduces a refinement over the original CESK machine, so we first prove it correct with respect to the concrete semantics, in a similar fashion to PDCFA.
%
\begin{theorem}[Correctness of stack frame refinement]\label{thm:refinement}
  The CES$\xi$K machine has a bisimulation with the CESK machine with fresh allocation.
\end{theorem}
This follows from the invariant that if a stack frame maps $\maddr$ to $\mval$, then the store maps $\maddr$ to $\mval$, meaning that the special lookup function is equivalent to lookup directly in the store (exploits immutability).
%
%That is, if we always allocated bindings in the heap.
%
%The additional escape information that allows us to elide allocation in the heap introduces a second level of reasoning, where the burden of proof falls on the escape analysis.
%
%The escape analysis adds information to use in the proof, that if $\mvar$ never escapes, then every reference to allocations of $\mvar$'s binding will always be able to use $\mframe$ for the lookup.
%
%With this proposition, we can prove that the CES$\xi$K machine that always allocates in the store has a bisimulation with the refined machine that exploits escape analysis to avoid unnecessary changes to the store.
%

@@

%% TODO: Cut here next.
There is contention among programming language researchers whether \rackett{call/cc} should be a language primitive\citep{ianjohnson:kiselyov:against-callcc}.
%
Alternative control operators have been proposed that delimit how much of the stack to capture, such as $\%$ (read ``prompt'') and capture operator ${\mathcal F}$ (read ``control'') \citep{ianjohnson:felleisen:control:1988}, or \texttt{reset} and \texttt{shift} \citep{ianjohnson:danvy:filinski:delim:1990}.
%
However, the stacks captured by these operators always extend the stack when invoked, rather than replace it like those captured with \rackett{call/cc}.
%
Continuations that have this extension behavior are called ``composable continuations.''
%
Stack extension poses a new challenge for pushdown analysis, since one application of a composable continuation means pushing an unbounded amount of frames onto the stack.
%
Vardoulakis' and Shivers' technique drops all knowledge of the stack at a continuation's invocation site; extension must preserve it.
%%

@@

The table-based semantics makes prompts a point of indirection, just like function calls.
%
Memoization also gets a new context to consider, calling a continuation, because composable continuations act like functions.
%
The new semantic spaces are then

@@

\autoref{fig:shift-reset-table0} has what one might naturally write using just this information.
%
\paragraph{Problem:} now $\rt$ continuations contain a store, and captured continuations can appear in the store.
%
This introduces an unbounded data structure that can cause the analysis to never terminate.

\paragraph{Solution:} we allow for a tunable approximation by using $\alloc$.
%
Captured continuations' store component in the $\rt$ context gets replaced with an address of possible stores that it would approximate.

@@

\section{Conclusion}
